---
title: "STATS141SL Final Project (Persian Eye)"
author: "Jiancong Qi, Zheng Wang, Janice Li, Joy Wang, Wenxin Zhou"
date: "2/19/2020"
output: html_document
---

```{r, message=FALSE}
library(readxl)
library(dplyr)
library(caret)
library(e1071)
library(corrplot)
library(ggplot2)
library(readr)
library(nlme)
```

### Data Cleaning
```{r, message=FALSE}
persian.UCLA = read_xlsx("/Users/jinyi/Downloads/PersianVCaucasianEyes.xlsx", sheet = 1)[,2:12]
persian.UCLA = na.omit(persian.UCLA)
persian.UCLA$Ethnicity = "Persian"

caucasian = read_xlsx("PersianVCaucasianEyes.xlsx", sheet = 2)[,2:12]
caucasian = na.omit(caucasian)

alldata = rbind(persian.UCLA, caucasian)

persian = read_xlsx("PersianVCaucasianEyes.xlsx", sheet = 3)[,2:9]
persian$Age = as.numeric(alldata %>% filter(Ethnicity == "Persian" & Sex == "M") %>% summarise(Age = median(Age)))
persian$Ethnicity = "Persian"

newAll = rbind(alldata[,c(11,2,5:10,1)], persian)
newTwo = alldata[,c(11,2,5:10,1)]
write.csv(newTwo,file="complete.csv")
newTwo$Ethnicity = as.factor(newTwo$Ethnicity)
newAll$Ethnicity = as.factor(newAll$Ethnicity)
persian$Ethnicity = as.factor(persian$Ethnicity)
male = newTwo %>% filter(Sex == "M")
female = newTwo %>% filter(Sex == "F")
```

###Exploratory Data Analysis


####Correlation of eye features
```{r, message=FALSE}
# All Data
corrplot(cor(newTwo[,c(1,3:8)]))

# Caucasian
corrplot(cor(newTwo[which(newTwo$Ethnicity=="Caucasian"),c(1,3:8)]))

# Persian
corrplot(cor(newTwo[which(newTwo$Ethnicity=="Persian"),c(1,3:8)]))

# Male
corrplot(cor(male[,c(1,3:8)]))

# Female
corrplot(cor(female[,c(1,3:8)]))
```

#####Observation:
1. MRD, PTB and TPS all show fairly high correlation between left and right eyes(~0.8).
2. TPS and PTB correlate with each other(~0.6). Such correlation is slightly higher among Caucasians and females.
2. Although not very strong, age negatively correlates with MRD(~0.4). We should consider adding age to the model as a covariate.


```{r}
#Investigate linear relation between age and MRD
ggplot(newTwo, aes(x=`MRD-1 (L)`, y=Age,fill=Ethnicity)) + geom_point(aes(color=Ethnicity)) + theme_classic()
```


####Distribution of Each Eye Feature
```{r}
ggplot(newAll, aes(x=`MRD-1 (L)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()
ggplot(newAll, aes(x=`MRD-1(R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()
ggplot(newAll, aes(x=`PTB (L)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()
ggplot(newAll, aes(x=`PTB (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()
ggplot(newAll, aes(x=`TPS (L)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()
ggplot(newAll, aes(x=`TPS (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()
```

#####Observation:
1. MRD shows Gaussian distribution, while TPS seems lognormal. PTB is ambiguous and may be either.
2. Caucasian and Persian have obviously different MRD on both left and right eyes, while TPS and PTB do not show significant difference among ethnic groups. Therefore we draw separate distribution plots for both genders below to investigate whether the difference was masked by the gender covariate.


####Distribution of Each Eye Feature, Partitioned by Gender
```{r}
ggplot(male, aes(x=`PTB (L)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("PTB (L) of Male") + theme(plot.title = element_text(hjust = 0.5))
ggplot(female, aes(x=`PTB (L)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("PTB (L) of Female") + theme(plot.title = element_text(hjust = 0.5))
ggplot(male, aes(x=`PTB (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("PTB (R) of Male") + theme(plot.title = element_text(hjust = 0.5))
ggplot(female, aes(x=`PTB (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("PTB (R) of Female") + theme(plot.title = element_text(hjust = 0.5))
ggplot(male, aes(x=`TPS (L)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("TPS (L) of Male") + theme(plot.title = element_text(hjust = 0.5))
ggplot(female, aes(x=`TPS (L)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("TPS (L) of Female") + theme(plot.title = element_text(hjust = 0.5))
ggplot(male, aes(x=`TPS (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("TPS (R) of Male") + theme(plot.title = element_text(hjust = 0.5))
ggplot(female, aes(x=`TPS (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("TPS (R) of Female") + theme(plot.title = element_text(hjust = 0.5))
```

#####Observation:
1. After splitting the data by gender, both PTB and TPS show obvious differences between the ethnic groups. Therefore, we should create separate models for both genders instead of neglecting the covariate.
2. The distribution of PTB(L) for Male looks abnormal. Since PTB(R), which is highly correlated to PTB(L), shows a normal Gaussian distribution, we assume that the exotic shape arised due to the small sample size(only 21 Caucasions and 13 Persians in this sample).


####Distribution of Variables Derived from Eye Features, Partitioned by Gender (Taking right eyes as examples)

```{r}
ggplot(male, aes(x=`MRD-1(R)`/`TPS (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("MRD-1(R)/TPS (R) of Male") + theme(plot.title = element_text(hjust = 0.5))
ggplot(female, aes(x=`MRD-1(R)`/`TPS (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("MRD-1(R)/TPS(R) of Female") + theme(plot.title = element_text(hjust = 0.5))
ggplot(male, aes(x=`MRD-1(R)`/`PTB (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("MRD-1(R)/PTB(R) of Male") + theme(plot.title = element_text(hjust = 0.5))
ggplot(female, aes(x=`MRD-1(R)`/`PTB (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("MRD-1(R)/PTB(R) of Female") + theme(plot.title = element_text(hjust = 0.5))
ggplot(male, aes(x=`PTB (R)`/`TPS (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("PTB(R)/TPS(R) of Male") + theme(plot.title = element_text(hjust = 0.5))
ggplot(female, aes(x=`PTB (R)`/`TPS (R)`, fill=Ethnicity)) + geom_density(alpha=.3)+ theme_classic()+ggtitle("PTB(R)/TPS(R) of Female") + theme(plot.title = element_text(hjust = 0.5))
```

```{r}
dat_male=data.frame(age=male$Age,'MRD-1(R)/TPS(R) of Male'=male$`MRD-1(R)`/male$`TPS (R)`,'MRD-1(R)/PTB(R) of Male'=male$`MRD-1(R)`/male$`PTB (R)`,'PTB(R)/TPS(R)'=male$`PTB (R)`/male$`TPS (R)`)
corrplot(cor(dat_male))
dat_female=data.frame(age=female$Age,'MRD-1(R)/TPS(R) of Male'=female$`MRD-1(R)`/female$`TPS (R)`,'MRD-1(R)/PTB(R) of Male'=female$`MRD-1(R)`/female$`PTB (R)`,'PTB(R)/TPS(R)'=female$`PTB (R)`/female$`TPS (R)`)
corrplot(cor(dat_female))
```


#####Observation:
1. According to the distribution plots, the derived variables, MRD/TPS and MRD/PTB are significantly differentiated by gender.
2. According to the correlation plot, males' ages correlate with their MRD/TPS and MRD/PTB, with coefficients of 0.34 and 0.5 respectively. Such correlations are weaker on females, but they still exist.

They above two observations suggest that we can potentially predict the two derived variables, MRD/TPS and MRD/PTB, with gender and age.


###CV Split
```{r, message=FALSE}
smp_siz = floor(0.8*nrow(newTwo))
set.seed(684323)
train_ind = sample(seq_len(nrow(newTwo)),size = smp_siz)
train = newTwo[train_ind,]
test = newTwo[-train_ind,]

write.csv(train,file="ann_train.csv")
write.csv(test,file="ann_test.csv")

smp_siz = floor(0.8*nrow(male))
train_ind = sample(seq_len(nrow(male)),size = smp_siz)
trainMale = male[train_ind,]
testMale = male[-train_ind,]

smp_siz = floor(0.8*nrow(female))
train_ind = sample(seq_len(nrow(female)),size = smp_siz)
trainFemale = female[train_ind,]
testFemale = female[-train_ind,]
```

###Data Modeling to Predict Ethnicity
```{r, message=FALSE}
# All
m1 = glm(Ethnicity ~ `MRD-1 (L)` + `PTB (L)` + Age + Sex, data = train, family = "binomial")

m2 = glm(Ethnicity ~ `MRD-1 (L)` * `TPS (L)` * `PTB (L)` + Age + Sex, data = train, family = "binomial")

m3 = glm(Ethnicity ~ `MRD-1 (L)` * `TPS (L)` * `PTB (L)` * Age * Sex, data = train, family = "binomial")

# Male/Female
m4 = glm(Ethnicity ~ `MRD-1 (L)` + `PTB (L)` + Age, data = trainFemale , family = "binomial")

m5 = glm(Ethnicity ~ `MRD-1 (L)` + `PTB (L)` + Age, data = trainMale , family = "binomial")
```


```{r, message=FALSE}
summary(m1)

summary(m2)

summary(m3)

summary(m4)

summary(m5)
```


```{r, message=FALSE}
confusionMatrix(table(predict(m1,test[,1:8],type = 'response') >=0.5,test$Ethnicity=="Persian"))[3]

confusionMatrix(table(predict(m2,test[,1:8],type = 'response') >=0.5,test$Ethnicity=="Persian"))[3]

confusionMatrix(table(predict(m3,test[,1:8],type = 'response') >=0.5,test$Ethnicity=="Persian"))[3]

confusionMatrix(table(predict(m4,testFemale[,1:8],type = 'response') >=0.5,testFemale$Ethnicity=="Persian"))[3]

confusionMatrix(table(predict(m5,testMale[,1:8],type = 'response') >=0.5,testMale$Ethnicity=="Persian"))[3]
```

```{bash}
# run the neural network as a reference to see performance
# of the logistic regression model
python FinalProj_ann.py
```

```{bash}
# run python script to prepare
# data for the next part
python reorganize.py
```


```{r,message=FALSE, warning=FALSE, paged.print = FALSE}
## run random effect model to see how the measures are differed by Ethnicity
random_effect_data <- read_csv("./random_effect_data.csv")

random_effect_MRD = lme(`MRD` ~ Ethnicity + Sex + Age, random_effect_data, random=~1|Person)
random_effect_PTB = lme(`PTB` ~ Ethnicity + Sex + Age, random_effect_data, random=~1|Person)
random_effect_TPS = lme(`TPS` ~ Ethnicity + Sex + Age, random_effect_data, random=~1|Person)
random_effect_MRD_TPS = lme(`MRD`/`TPS` ~ Ethnicity + Sex + Age, random_effect_data, random=~1|Person)
random_effect_MRD_PTB = lme(`MRD`/`PTB` ~ Ethnicity + Sex + Age, random_effect_data, random=~1|Person)
random_effect_PTB_TPS = lme(`PTB`/`TPS` ~ Ethnicity + Sex + Age, random_effect_data, random=~1|Person)

summary(random_effect_MRD)
summary(random_effect_PTB)
summary(random_effect_TPS)
summary(random_effect_MRD_TPS)
summary(random_effect_MRD_PTB)
summary(random_effect_PTB_TPS)
```

